---
title: "Untitled"
author: "Kameni"
date: "Sunday, September 21, 2014"
output: html_document
---
# Title: 
(Your document should have a title that briefly summarizes your data analysis)


## synopsis
(Synopsis: Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences)


## Data Loading and  Processing
```{r}
Sys.setenv(LANG = "en_US.UTF-8")

```


 We start the anlysis by downloading the data.
```{r}

if (!file.exists("./StormData")) {dir.create("./StormData")}
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
#download.file(url,"./StormData/dataSet.zip")


```



####read in the data. Since the data set 
```{r}
library(R.utils)
if (!file.exists("./StormData/dataSet.csv")) {
    bunzip2 ("./StormData/dataSet.zip", "./StormData/dataSet.csv")
}

stormData<- read.csv ("./StormData/dataSet.csv",stringsAsFactors=F)
head(stormData)
str(stormData)
```




We will start by reducing the data set and keep only the columns we need for our analysis.

 We take a look on the column names
```{r}
columNames<-names(stormData)

columNames
```



 We select the relavant columns
```{r}
selectedColumns<-columNames[c(2,8,23,24,25,26,27,28)] 

selectedColumns

```



 we reduce the dimensions of the data set
```{r}
reducedDf<-stormData[,selectedColumns]
dim(reducedDf)
```

#### Processing the the BGN_DATE column

```{r}
# remove the "0:00:0" time indication to make the BGN_DATE column easy to handle
reducedDf$BGN_DATE<-gsub(" 0:00:00","",reducedDf$BGN_DATE)

# replace the / by the - 
reducedDf$BGN_DATE<-gsub("/","-",reducedDf$BGN_DATE)


reducedDf$BGN_DATE<-as.Date(reducedDf$BGN_DATE,format = "%m-%d-%Y")

head(reducedDf$BGN_DATE)

```

 
 
 We will replace the "BGN_DATE"" column by a new column with only the year whitout month an day. This step will help to easily reduce the data set by removing observations before 1996.
 
 
```{r}
#select the 4 first characters of the BGN_DATE column
#the 4 first characters correspond to the year
years<-as.numeric(unclass(substr(reducedDf$BGN_DATE, 1, 4)))

#Add the column  "years""
reducedDf<-cbind(years,reducedDf) # add the new column

#remove the "BGN_DATE"column from data set
reducedDf$BGN_DATE <-NULL

#view the data 
dim(reducedDf)
head(reducedDf)

```


We reduce the data frame: We will concentrate our analysis on the  observations between 1996 and 2011
```{r}
yearsDf<- reducedDf[reducedDf$years >= 1996,]
head(yearsDf)
dim(yearsDf)
```


##### processing the Columns: FATALITIES & INJURIES

We remove all rows where either the pair of fatalities and injuries are both zero, or propertydamage and cropdamage both are zero. 
If a row has PROPDMG = 0, CROPDMG=0, INJURIES+FATALITIES=0 then that observation has no use in this project.  

Subsetting for only the observations where one of those variables is >0 
```{r}
rmZero<-yearsDf[(yearsDf$FATALITIES == 0 & yearsDf$INJURIES)|yearsDf$PROPDMG  > 0|yearsDf$CROPDMG > 0,]

dim(rmZero)

```



#### Processing the column "EVTYPE"
 We want to see the  different event types that are recorded in the data set
 
 
```{r}
      Event<- unique(tolower(rmZero$EVTYPE)) # create a vector of different Event type

```

there are 985 different Event types...
```{r}
length(Event)
```

the list of these Event types shows some anomalies.
```{r}
Event
```

  
 We want to  remove white space at the begining of some rows. We have to repeat
 the operation several times since some rows have more than one white spaces at the begining.
  
  
```{r}
#  find and remove space before the Event name 
for(i in 1:4) {   # repeat the action  times
Event<-gsub("^ ", "",Event)
}

```


### Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?



 
 



### Across the United States, which types of events have the greatest economic consequences?
•1 Check for outliers, high leverage, overdispersion
•2 Estimate degrees of freedom to adjust for unmeasured confounders
•3 Sensitivity analysis wrt


## Results 

```{r}
sessionInfo()
```
